---
title: 'CUNY MSDS Data 605 HW #8'
author: "Samantha Deokinanan"
date: "20th October, 2019"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '4'
---
***
### Problem Set #1
A company buys 100 light bulbs, each of which has an exponential lifetime of 1000 hours. What is the expected time for the first of these bulbs to burn out?

This exponential density has mean = $\frac {\mu}{n}$, according to question 10.

Thus, if $X_{i}$ is an independent random variable, $E[X_{i}] = \frac {1}{\lambda_{i}} = 1000$. Therefore, $E[min X_{i}] = \frac {1000}{100} = 10$ hours is expected to pass before the first bulb burn out.

***

### Problem Set #2
Assume that $X_{1}$ and $X_{2}$ are independent random variables, each having an exponential density with parameter $\lambda$. Show that $Z = X_{1} - X_{2}$ has density ${f}_{n}(z)=\frac {1}{(n-1)!} {(-log z)}^{n-1}$.

The probability distribution of the sum of two or more independent random variables is the convolution of their individual distributions. The general formula for the distribution of the sum Z = X + Y of two independent integer-valued and discrete random variables is

$P(Z=z)=\sum _{k=-\infty}^{\infty}{P(X=k)P(Y=z-k)}$ 

Let $Y = -X_{2}$, then the counterpart for independent continuously distributed random variables with density functions $f_{X_{1}}, f_{X_{2}}$ is 

$=\int _{-\infty}^{\infty}{{f}_{X_{1}}({x}_{1}){f}_{X_{2}}({x}_{2})dx}$

$=\int _{-\infty}^{\infty}{{f}_{{X}_{1}}({x}_{1}){f}_{ {-X}_{2}}(z-{x}_{1})dx}$

$=\int _{-\infty}^{\infty}{{f}_{ {X}_{1}}({x}_{1}){f}_{{X}_{2}}({x}_{1}-z)dx_{1}}$ 

Now for z < 0, the above can be expressed as the exponential distribution ${f}_{Z}(z)$

$=\int _{0}^{\infty}{\lambda{e}^{-\lambda{x}_{1}}\lambda {e}^{-\lambda {(x}_{1}-z)}d{x}_{1}}$

$=\int _{0}^{\infty }{{\lambda}^{2}{e}^{-2\lambda {x}_{1}}{e}^{\lambda z) }d{x}_{1}}$

$= \lambda {e}^{\lambda z}(-\frac {1}{2} {e }^{-2\lambda {x}_{1}}\overset {\infty}{\underset {0}{|})}$ 

$=\frac {1}{2} \lambda {e}^{\lambda z}$

The question ask for $Z = X_{1} - X_{2}$, but note that $-Z = X_{2} - X_{1}$ and $X_{i}$ are iid, thus Z and -Z have the same distribution that is symmetric around 0, ie ${f}_{Z}(z)={f}_{-Z}(-z)$.

Therefore, for $z \geq 0$, ${f}_{Z}(z)=\frac {1}{2} \lambda {e}^{\lambda (-z)}$. 

And altogether, ${f}_{Z}(z)=\frac {1}{2} \lambda {e}^{-\lambda |z|}$.


***

### Problem Set #3

Let $X$ be a continuous random variable with mean $\mu$ = 10 and variance $\sigma^{2}$ = $\frac {100}{3}$. Using Chebyshev's Inequality, find an upper bound for the following probabilities.

(a) P(|X - 10| $\geq$ 2)
```{r}
k <- 2/sqrt(100/3)
1/k^2
```
So, the upper bound is 1.

(b) P(|X - 10| $\geq$ 5)
```{r}
k <- 5/sqrt(100/3)
1/k^2
```
So, the upper bound is 1.

(c) P(|X - 10| $\geq$ 9)
```{r}
k <- 9/sqrt(100/3)
1/k^2
```
So, the upper bound is 0.412.

(d) P(|X - 10| $\geq$ 20)
```{r}
k <- 20/sqrt(100/3)
1/k^2
```
So, the upper bound is 0.083.

***