---
title: 'CUNY MSDS Data 605 HW #3'
author: "Samantha Deokinanan"
date: "15th September, 2019"
output:
  pdf_document:
    toc: yes
    toc_depth: '4'
---
***

### Problem Set #1.1
What is the rank of matrix $A=\begin{bmatrix} 1 & 2 & 3 & 4 \\ -1 & 0 & 1 & 3 \\ 0 & 1 & -2 & 1 \\ 5 & 4 & -2 & -3 \end{bmatrix}$?

```{r}
# The rank of a matrix is the number of pivots it has.
# A few row reduction operations to obtain the pivots 
a = matrix(c(1,-1,0,5,2,0,1,4,3,1,-2,-2,4,3,1,-3), ncol = 4)
a[2,] = a[1,] + a[2,]
a[4,] = -5*a[1,] + a[4,]
a[2,] = -2*a[3,] + a[2,]
temp = a[2,]
a[2,] = a[3,]
a[3,] = temp
a[4,] = 6*a[2,] + a[4,]
a[4,] = (29/8)*a[3,] + a[4,]

# The poviting positions show that matrix A is rank 4.
a
```

***
### Problem Set #1.2
Given an $m \times n$ matrix where $m > n$, what can be the maximum rank? The minimum rank, assuming that the matrix is non-zero?

An $m \times n$ matrix where $m > n$ may have maximum rank of $n$, and the minimum rank, assuming that the matrix is non-zero, of 1.

***
### Problem Set #1.3
What is the rank of matrix $B=\begin{bmatrix} 1 & 2 & 1 \\ 3 & 6 & 3 \\ 2 & 4 & 2 \end{bmatrix}$?

Upon visual inspection, it is clear that row 2 and 3 of matrix B are scalar multiples of row 1; that is, row 2 and 3 are equal to trice and twice row 1, respectively. Therefore, the rows are linearly dependent, making its rank = 1.

***
### Problem Set #2
Compute the eigenvalues and eigenvectors of the matrix $A=\begin{bmatrix} 1 & 2 & 3 \\ 0 & 4 & 5 \\ 0 & 0 & 6 \end{bmatrix}$.

Solve $det(A - \lambda I)=0$, where $\lambda$ is the eigenvalue. This gives $det\left( \begin{bmatrix} 1 & 2 & 3 \\ 0 & 4 & 5 \\ 0 & 0 & 6 \end{bmatrix}-\lambda \begin{bmatrix} 1 & 0 & 0 \\ 0 & 1 & 0 \\ 0 & 0 & 1 \end{bmatrix} \right) =0$ that is, $det\left( \begin{bmatrix} 1-\lambda  & 2 & 3 \\ 0 & 4-\lambda  & 5 \\ 0 & 0 & 6-\lambda  \end{bmatrix} \right) =0$ which conveniently reduces to the characteristic polynomial $(\lambda -1)(\lambda -4)(\lambda -6)$. 

Thus, the eigenvalues are ${ \lambda  }_{ 1 }=1, { \lambda  }_{ 2 }=4,$ and ${ \lambda  }_{ 3 }=6.$

***

Now, the eigenvector, $v$, for a specific eigenvalue is such that $(A - \lambda I)v = 0$.

So, for ${\lambda  }_{ 1 }=1$, it is reduced to $\left( \begin{bmatrix} 0 & 2 & 3 \\ 0 & 3 & 5 \\ 0 & 0 & 5 \end{bmatrix} \right) \left[ \begin{matrix} { v }_{ 1 } \\ { v }_{ 2 } \\ { v }_{ 3 } \end{matrix} \right] =0$ and solving the following system of equations gives the eigenvector.

```{r warning=FALSE}
lambda_1 = matrix(c(0,0,0,2,3,0,3,5,5), ncol = 3)
library(matlib)
Solve(lambda_1)
```

Note that $v_{1}$ did not show up in the equations, this means that $\left[ \begin{matrix} 0 & 1 & 0 \\ 0 & 0 & 0 \\ 0 & 0 & 1 \end{matrix} \right] \left[ \begin{matrix} { v }_{ 1 } \\ 0 \\ 0 \end{matrix} \right] =\left[ \begin{matrix} 0 \\ 0 \\ 0 \end{matrix} \right]$ regardless of the value of $v_{1}$. Thus, the eigenvector associated with ${\lambda  }_{ 1 }=1$ is $\left[ \begin{matrix} 1 \\ 0 \\ 0 \end{matrix} \right]$.

For ${\lambda  }_{ 2 }=4$, the equation is reduced to $\left( \begin{bmatrix} -3 & 2 & 3 \\ 0 & 0 & 5 \\ 0 & 0 & 2 \end{bmatrix} \right) \left[ \begin{matrix} { v }_{ 1 } \\ { v }_{ 2 } \\ { v }_{ 3 } \end{matrix} \right] =0$.

```{r}
lambda_4 = matrix(c(-3,0,0,2,0,0,3,5,2), ncol = 3)
Solve(lambda_4)
```

The system of equations are $\left\{ \begin{matrix} { v }_{ 1 }=\frac { 2 }{ 3 } { v }_{ 2 } \\ { v }_{ 3 }=0 \end{matrix} \right\}$, and if $v_{2} = 3$, the eigenvector associated with ${\lambda  }_{ 2 }=4$ is $\left[ \begin{matrix} 2 \\ 3 \\ 0 \end{matrix} \right]$.

Lastly, for ${\lambda  }_{ 3 }=6$, the equation is reduced to $\left( \begin{bmatrix} -5 & 2 & 3 \\ 0 & -2 & 5 \\ 0 & 0 & 0 \end{bmatrix} \right) \left[ \begin{matrix} { v }_{ 1 } \\ { v }_{ 2 } \\ { v }_{ 3 } \end{matrix} \right] =0$.

```{r}
lambda_6 = matrix(c(-5,0,0,2,-2,0,3,5,0), ncol = 3)
Solve(lambda_6)
```

The system of equations are $\left\{ \begin{matrix} { v }_{ 1 }=\frac { 8 }{ 5 } { v }_{ 3 } \\ { v }_{ 2 }=\frac { 5 }{ 2 } { v }_{ 3 } \end{matrix} \right\}$, and if $v_{3} = 10$, the eigenvector associated with ${\lambda  }_{ 3 }=6$ is $\left[ \begin{matrix} 16 \\ 25 \\ 10 \end{matrix} \right]$.

Altogether, the eigenvalues and eigenvectors for matrix $A=\begin{bmatrix} 1 & 2 & 3 \\ 0 & 4 & 5 \\ 0 & 0 & 6 \end{bmatrix}$ are $\lambda =\left[ \begin{matrix} 1 \\ 4 \\ 6 \end{matrix} \right]$, and $v=\left[ \begin{matrix} 1 & 2 & 16 \\ 0 & 3 & 25 \\ 0 & 0 & 10 \end{matrix} \right]$.

***
This test confirmation ensures that $(A - \lambda I)v=0$ for each eigenvalue and its associated eigenvector.

```{r}
# Test confirmation
A = matrix(c(1,0,0,2,4,0,3,5,6), ncol = 3)

# For lambda = 1
eigval = diag(1,3)
eigvec = matrix(c(1,0,0), ncol = 1)
(A-eigval)%*%eigvec

# For lambda = 4
eigval = diag(4,3)
eigvec = matrix(c(2,3,0), ncol = 1)
(A-eigval)%*%eigvec

# For lambda = 6
eigval = diag(6,3)
eigvec = matrix(c(16,25,10), ncol = 1)
(A-eigval)%*%eigvec
```

***
### Works Cited
* Geezer, Robert A. "A first course in linear algebra." 2008.