---
title: "CUNY MSDS Data 605 HW #1"
author: "Samantha Deokinanan"
date: "1st September, 2019"
output:
  html_document:
    theme: journal
    toc: TRUE
    toc_depth: 4
---
***

### Problem Set #1 

You can think of vectors representing many dimensions of related information. For instance, Netflix might store all the ratings a user gives to movies in a vector. This is clearly a vector of very large dimensions (in the millions) and very sparse as the user might have rated only a few movies. Similarly, Amazon might store the items purchased by a user in a vector, with each slot or dimension representing a unique product and the value of the slot, the number of such items the user bought. One task that is frequently done in these settings is to find similarities between users. And, we can use dot-product between vectors to do just that. As you know, the dot-product is proportional to the length of two vectors and to the angle between them. In fact, the dot-product between two vectors, normalized by their lengths is called as the cosine distance and is frequently used in recommendation engines.

***

#### Q1
Calculate the dot product *u.v* where *u* = [0.5; 0.5] and *v* = [3; -4].

```{r}
u <- c(0.5,0.5)
v <- c(3,-4)

dot_uv <- u%*%v
```
```{r echo=FALSE}
sprintf("The dot product of u.v is %.1f.", dot_uv)
```

***

#### Q2
What are the lengths of *u* and *v*? 

The lengths of a vector is calculated by $\sqrt { \sum _{ i=1 }^{ m }{ { \left| { \left[ u \right]  }_{ i } \right|  }^{ 2 }}}$.

```{r}
length_u <- sqrt(sum(u^2))
length_v <- sqrt(sum(v^2))
```
```{r echo=FALSE}
sprintf("The lengths of u and v are %.2f and %.1f, respectively.", length_u, length_v)
```

***

#### Q3
What is the linear combination: 3*u* - 2*v*?

```{r}
lc <- 3*u - 2*v
```
```{r echo=FALSE}
sprintf("The linear combination of 3u-2v is [%.1f; %.1f]", lc[1], lc[2])
```

***

#### Q4
What is the angle between *u* and *v*?

We know that the inner product of two vectors is $\left<u,\quad v\right>\quad=\quad\left\|u\right\|\quad*\quad\left\|v \right\|\quad*\quad\cos{(\alpha)}$, where $\alpha$ is the angle between the two vectors. Therefore, $\alpha$ can be found using $\alpha \quad =\quad\arccos{\left(\frac{\left<u,\quad v\right>}{\left\|u\right\|*\left\|v\right\|}\right)}$.


```{r}
alpha <- acos(dot_uv / (length_u * length_v)) * (180 / pi)
```
```{r echo=FALSE}
sprintf("The angle between u and v is %.2f degrees.", alpha)
```

***

### Problem Set #2 

Set up a system of equations with 3 variables and 3 constraints and solve for x. Please write a function in R that will take two variables (matrix A & constraint vector b) and solve using elimination. Your function should produce the right answer for the system of equations for any 3-variable, 3-equation system. The approach that you should employ is to construct an Upper Triangular Matrix and then back-substitute to get the solution. Alternatively, you can augment the matrix A with vector b and jointly apply the Gauss Jordan elimination procedure.

$\left[ \begin{matrix} 1 & 1 & 3 \\ 2 & -1 & 5 \\ -1 & -2 & 4 \end{matrix} \right] \left[ \begin{matrix} { x }_{ 1 } \\ { x }_{ 2 } \\ { x }_{ 3 } \end{matrix} \right] =\left[ \begin{matrix} 1 \\ 2 \\ 6 \end{matrix} \right]$ 

***

#### function Solve4X

Solve4X is a function that takes two variables (matrix A & constraint vector b) and solve for x using Gauss Jordan elimination method for any 3-variable, 3-equation system.
```{r}
Solve4X <- function(A,b){
  # Write the augmented matrix of the system
  aug_mat <- cbind(A,b)
  
  # Get matrix dimension
  m <- nrow(aug_mat)
  n <- ncol(aug_mat)
  
  # The pivot process to obtain reduced row echelon form
  p <- 1 # initating pivot position
  
  for(k in 1:m) {
    if (n <= p) 
      break;
    i <- k
    # Get pivot of a row
    while(aug_mat[i,p] == 0) {
      i <- i + 1
      if (m == i) {
        i <- k
        p <- p + 1
        if (n == p) 
          return(aug_mat)
      }
    }
    # Swap rows with another were pivot is non-zero
    swap <- aug_mat[i,]
    aug_mat[i,] <- aug_mat[k,]
    aug_mat[k,] <- swap
    
    #  Divide the k-th row by kp-th entry to make the pivot = 1
    aug_mat[k,] <- aug_mat[k,] / aug_mat[k,p]
    
    # Eliminate all other entries in the p-th column by subtracting suitable multiples of the i-th row from the other rows
    for(i in 1:m) {
      if (i != k)
        aug_mat[i,] <- aug_mat[i,] - aug_mat[k,] * aug_mat[i,p]
    }
    # Increase pivot position by 1 to choose the new pivot, and repeat
    p <- p + 1
  }
  return(aug_mat)
}
```
```{r}
A <- matrix(c(1,2,-1,1,-1,-2,3,5,4), nrow=3)
b <- c(1,2,6)

Solve4X(A,b)
```

### Works Cited

* Beezer, Robert A. "A first course in linear algebra." 2008: pg 28-32.