---
title: 'CUNY MSDS Data 605 HW #9'
author: "Samantha Deokinanan"
date: "27th October, 2019"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '4'
---
***
### Problem Set #1
The price of one share of stock in the Pilsdorff Beer Company (see Exercise 8.2.12) is given by $Y_{n}$ on the nth day of the year. Finn observes that the differences $X_{n} = Y_{n+1} − Y_{n}$ appear to be independent random variables with a common distribution having mean $\mu$ = 0 and variance $\sigma^{2} = \frac {1}{4}$. If $Y_{1}$ = 100, estimate the probability that $Y_{365}$ is  
(a) $\geq$ 100.  
(b) $\geq$ 110.  
(c) $\geq$ 120.  

##### Solution
$E[X_{n}]=E[Y_{n+1} - Y_{1}] = E[Y_{n+1}] - E[Y_{n}]$

With $E[X_{i}] = 0$, $E[Y_{n+1}] = E[Y_{n}]$, and $E[Y_{1}] = 100$

$Var[Y_{365}] = Var[Y_{364} + X_{365}]$  
$=Var[Y_{364}]+Var[X_{365}]$  
$=Var[Y_{364}]+\sigma$  
$=Var[Y_{363}]+2\sigma=...=365\sigma$  
$=\frac{365}{4}$  

(a) $P(Y_{365} \geq 100)$  
Since $E[Y_{365}] = 100$, $P(Y_{365} \geq 100) = 0.5$.

(b) $P(Y_{365} \geq 110)$  
Using the Central Limit Theorem, we can standardized $P(Y_{365} \geq 110) \rightarrow P(\frac {Y_{n} - Y_{1}}{\sigma\sqrt{n-1}} \geq 110) = P(z \geq \frac {10}{0.5 \sqrt{364}})$

```{r}
1-pnorm((110-100)/(0.5*sqrt(364)))
```

(c) $P(Y_{365} \geq 120)$  
Using the Central Limit Theorem, we can standardized $P(Y_{365} \geq 120) \rightarrow P(\frac {Y_{n} - Y_{1}}{\sigma\sqrt{n-1}} \geq 120) = P(z \geq \frac {20}{0.5 \sqrt{364}})$

```{r}
1-pnorm((120-100)/(0.5*sqrt(364)))
```

***

### Problem Set #2
Calculate the expected value and variance of the binomial distribution using the moment generating function.

##### Solution

The probability mass function is $f(x) = C(n,x)p^{x}(1 - p)^{n - x}$.

Using the pmf, the moment generating function of x is:  
$M(t)=\sum _{x=0}^{n}{{e}^{tx}C(n,x){p}^{x}}{(1-p)}^{n-x}$

Combining the exponent of x and using the binominal formula, $M(t)=[(1-p)+pe^{t}]^{n}$.

The expected value can be calculated as follows:  
$E[X]=M^{\prime}(t) \rightarrow$ the first derivative.  
$=npe^{t}[(1-p)+pe^{t}]^{n-1}$

Evaluating at t=0, $M'(0)=npe^{0}[(1-p)+pe^{0}]^{n-1}=np$.

The variance can be calculated as follows:  
$Var[X]= E[X^{2}]-(E[X]^{2})$  

We therefore need the second derivative, $M"(t) \rightarrow$
$=npe^{t}((n-1)((1-p)+pe^{t})^{n-2}pe^{t})+((1-p)+pe^{t})^{n-1}(npe^{t})$ 
$=npe^{t}((1-p)+pe^{t})^{n-2}((1-p)+npe^{t})$

Evaluating at t=0,  
$E[X^{2}]=M^{\prime\prime}(0)=np((1-p)+p)^{n-2}((1-p)+np)=np((1-p)+np)$.

Now, $Var[X]= E[X^{2}]-(E[X]^{2})$  
$= np((1-p)+np) - (np)^{2}=np(1-p)$

Therefore, $E[X]=np$ and $Var[X]=np(1-p)$.

***

### Problem Set #3
Calculate the expected value and variance of the exponential distribution using the moment generating function.

##### Solution

The probability density function is ${f}_{X}(x)=\begin{cases} \lambda {e}^{-\lambda x} \\ 0 \end{cases}\begin{matrix} x \in [0,\infty) \\ x \notin [0,\infty) \end{matrix}$.

Using the pdf, the moment generating function of x is:
$M_{X}(t)=\int _{-\infty}^{\infty}{{e}^{tx}{f}_{X}(x)dx}$  
$=\int _{0}^{\infty}{{e}^{tx}\lambda {e}^{-\lambda x}dx}$  
$=\lambda \int _{0}^{\infty}{{e}^{((t-\lambda)x)}dx}$  
$=\lambda \left[\frac {1}{t-\lambda}{e}^{((t-\lambda )x)}\right] ^{\infty}_{0}$  
$=\lambda \left[0-\frac {1}{t-\lambda}\right]$  
$=\frac {\lambda}{\lambda -t}$  

The expected value can be calculated as follows:  
$E[X]=M^{\prime}_{X}(t) \rightarrow$ the first derivative.  
$=\frac {\lambda}{(\lambda -t)^{2}}$

Evaluating at t=0,
$M^{\prime}_{X}(0)=\frac {\lambda}{(\lambda -0)^{2}}=\frac {1}{\lambda}$.

The variance can be calculated as follows:  
$Var[X]= E[X^{2}]-(E[X]^{2})$  

We therefore need the second derivative, $M"(t) \rightarrow$  
$=\frac {2\lambda}{(\lambda -t)^{3}}$

Evaluating at t=0, 
$M^{\prime\prime}(0)=\frac {2\lambda}{(\lambda -0)^{3}}=\frac {2}{\lambda ^{2}}$.

Now, $Var[X]= E[X^{2}]-(E[X]^{2}) = \frac {2}{\lambda ^{2}} - \frac {1}{\lambda ^{2}} = \frac {1}{\lambda ^{2}}$

Therefore, $E[X]=\frac {1}{\lambda}$ and $Var[X]=\frac {1}{\lambda ^{2}}$.

***