---
title: 'CUNY MSDS Data 605 HW #5'
author: "Samantha Deokinanan"
date: "29th September, 2019"
output:
  html_document:
    theme: journal
    toc: TRUE
    toc_depth: 4
---
***
### Problem Set #1
Choose independently two numbers B and C at random from the interval [0, 1] with uniform density. Prove that B and C are proper probability distributions. Note that the point (B,C) is then chosen at random in the unit square.

A probability distribution function on x, *p(x)*, is a measure of the likelihood of observing x over some range, $x_{min} < x < x_{max}$. *p(x)* must be normalizable, such that $\int _{ { x }_{ min } }^{ { x }_{ max } }{ dx\quad p(x) = 1 }$.

```{r}
# Create two random variables, B and C, by sampling n numbers from [0,1]
n = 50000
B1 = runif(n, min = 0, max = 1)
B = data.frame(B1)
C1 = runif(n, min = 0, max = 1)
C = data.frame(C1)

# Plots of the density estimate and finding the area under the curve will give the probability of x
library(ggplot2)
plot1 = ggplot(B, aes(x = B1)) + geom_density() + labs(title = "Density Plot of B", y = "p(x)", x = "x")
plot2 = ggplot(C, aes(x = C1)) + geom_density() +  labs(title = "Density Plot of C", y = "p(x)", x = "x")
gridExtra::grid.arrange(plot1, plot2, ncol=2)
```

***

### Problem Set #2
Find the probability that:

(a) B + C < 1/2

```{r}
# Density plot representation
BplusC = B + C
ggplot(BplusC, aes(x = B1)) +  geom_density() +  labs(title = "Density Plot of B + C < 1/2", y = "p(x)", x = "x")

# Calculation of the probability
count = 0
for(i in 1:n){
  if(B1[i] + C1[i] < 0.5){
    count = count + 1
  }
}
print(sprintf("The probability that B + C < 1/2 is %.03f", count/n))
```

***
(b) BC < 1/2
```{r}
# Density plot representation
BtimesC = B * C
ggplot(BtimesC, aes(x = B1)) +  geom_density() +  labs(title = "Density Plot of BC < 1/2", y = "p(x)", x = "x")

# Calculation of the probability
count = 0
for(i in 1:n){
  if(B1[i] * C1[i] < 0.5){
    count = count + 1
  }
}
print(sprintf("The probability that BC < 1/2 is %.03f", count/n))
```

***
(c) |B - C| < 1/2
```{r}
# Density plot representation
absBC = abs(B - C)
ggplot(absBC, aes(x = B1)) +  geom_density() +  labs(title = "Density Plot of |B - C| < 1/2", y = "p(x)", x = "x")

# Calculation of the probability
count = 0
for(i in 1:n){
  if(abs(B1[i] - C1[i]) < 0.5){
    count = count + 1
  }
}
print(sprintf("The probability that |B − C| < 1/2 is %.03f", count/n))
```

***
(d) max{B,C} < 1/2
```{r}
# Density plot representation
A = 0
maxBC = cbind(B1, C1)
for (i in 1:n){
  A[i] = max(maxBC[i,])
}
A = data.frame(A)
ggplot(A, aes(x = A)) +  geom_density() +  labs(title = "Density Plot of max{B,C} < 1/2", y = "p(x)", x = "x")

# Calculation of the probability
count = 0
for(i in 1:n){
  if(max(B1[i],C1[i]) < 0.5){
    count = count + 1
  }
}
print(sprintf("The probability that max{B,C} < 1/2 is %.03f", count/n))
```

***
(e) min{B,C} < 1/2
```{r}
# Density plot representation
A = 0
minBC = cbind(B1, C1)
for (i in 1:n){
  A[i] = min(minBC[i,])
}
A = data.frame(A)
ggplot(A, aes(x = A)) +  geom_density() +  labs(title = "Density Plot of min{B,C} < 1/2", y = "p(x)", x = "x")

# Calculation of the probability
count = 0
for(i in 1:n){
  if(min(B1[i],C1[i]) < 0.5){
    count = count + 1
  }
}
print(sprintf("The probability that min{B,C} < 1/2 is %.03f", count/n))
```

***
