---
title: 'CUNY MSDS Data 605 HW #2'
author: "Samantha Deokinanan"
date: "8th September, 2019"
output:
  pdf_document:
    toc: yes
    toc_depth: '4'
---
***

### Problem Set #1.1 
Show that ${ A }^{ T }A \neq A{ A }^{ T }$ in general. (Proof and demonstration)

#### Proof
Let $A \in { R }^{ m\times n }$, in general, there are two cases. These are:

##### Case 1
In most cases, when $m=n$, ${ A }=\left[ \begin{matrix} a & b \\ c & d \end{matrix} \right]$, and ${ A }^{ T }=\left[ \begin{matrix} a & c \\ b & d \end{matrix} \right]$,  then 

${ A }^{ T }A = \begin{bmatrix} a & c \\ b & d \end{bmatrix}\begin{bmatrix} a & b \\ c & d \end{bmatrix}=\begin{bmatrix} { a }^{ 2 }+{ c }^{ 2 } & ab+cd \\ ab+dc & { b }^{ 2 }+{ d }^{ 2 } \end{bmatrix}$

$A{ A }^{ T } = \begin{bmatrix} a & b \\ c & d \end{bmatrix}\begin{bmatrix} a & c \\ b & d \end{bmatrix}=\begin{bmatrix} { a }^{ 2 }+{ b }^{ 2 } & ac+bd \\ ac+bd & { c }^{ 2 }+{ d }^{ 2 } \end{bmatrix}$

Since matrix multiplication is not commutative, ${ A }^{ T }A \neq A{ A }^{ T }$.

##### Case 2
If $m \neq n$, $A$ is size $m \times n$, and ${ A }^{ T }$ will be an $n \times m$ matrix. Then ${ A }^{ T }A$ will be a square matrix of size $n \times n$ and $A{ A }^{ T }$ will be a square matrix of size $m \times m$, making these two matrices not equal to each other. This will hold for any non-square matrix $A$. 

#### Demonstration

```{r}
# Let A be an nxn matrix 
set.seed(525)
A <- matrix(runif(2*2), ncol = 2)
A

# Its transpose
At <- t(A)
At

# A^T * A will give
AtA <- At %*% A
AtA

# AA^T will give
AAt <- A %*% At
AAt

# Test if they are equal
AtA == AAt
```

***

### Problem Set #1.2
For a special type of square matrix $A$, we get ${ A }^{ T }A =  A{ A }^{ T }$. Under what conditions could this be true? (Hint: The Identity matrix $I$ is an example of such a matrix).

The conditions that may lead to ${ A }^{ T }A = A{ A }^{ T }$ are:

1. When $A$ is a symmetric matrix, i.e. $A = {A}^{T}$, then ${ A }^{ T }A = A{ A }^{ T } \rightarrow AA=AA$
2. When $A$ is an orthogonal matrix, i.e. ${ A }^{ T } = { A }^{ -1 }$, then ${ A }^{ T }A =  A{ A }^{ T } \rightarrow { A }^{ -1 }A =  A{ A }^{ -1 } \rightarrow I = I$

***

### Problem Set #2 
Write an R function to factorize a square matrix A using Lower-Upper or Lower-Diagonal-Upper factorization, whichever is prefer.

#### function factorize
*factorize* is a function that factorize a square matrix A into its Lower-Upper triangular matrices such that $A = LU$, where L is a unit lower triangular matrix and U is the echelon form of A.

```{r}
factorize <- function(A){
  # Ensure A is n by n
  size <- dim(A)
  if (size[1] != size[2]){
    stop('A is not n by n')
  } 
  # number of columns
  nc <- size[1]
  # Initialize L as an identity matrix and U as matrix A 
  L <- diag(nc)
  U <- A
  # Iterate operations until nc - 1 column
    for (i in 1:(nc-1)){
  # Reduce matrix A to an echelon form U by a sequence of row replacement operations 
        j <- i + 1
        for (j in (i + 1):nc){
            alpha <- -U[j,i] / U[i,i]
            U[j,i]  <- 0
            i2 <- i + 1
            for (i2 in (i + 1):nc){
                U[j,i2] <- U[j,i2] + alpha * U[i,i2] 
            }
  # Place entries in L such that the same sequence of row operations reduces L to I
            L[j,i] <- -alpha
        } 
    }
    result <- list("Lower Triangular Matrix" = L, "Upper Triangular Matrix" = U)
    return(result)
}
```

***

#### Demonstration

```{r}
A <- matrix(c(3,-3,6,-9,-7,5,-4,5,-2,1,0,-5,2,0,-5,12), nrow=4)
A
LU <- factorize(A)
LU

# Test confirmation
L <- LU[[1]]
U <- LU[[2]]
A == L%*%U
```

***

### Works Cited
* Geezer, Robert A. "A first course in linear algebra." 2008.