---
title: 'CUNY MSDS Data 605 HW #15'
author: "Samantha Deokinanan"
date: "8th December, 2019"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '4'
---
***  
### Problem 1  

Find the equation of the regression line for the given points.

```{r}
points = data.frame(x = c(5.6,6.3,7,7.7,8.4), y = c(8.8,12.4,14.8,18.2,20.8))
model = lm(points$y ~ points$x)
summary(model)
```

The equation of the regression line is $f(x) = 4.26x - 14.8$.

***
### Problem 2  

Find all local maxima, local minima, and saddle points for the function given below. Write your answer(s) in the form (x,y,z). Separate multiple points with a comma.

$$\begin{align} 
f(x,y) = 24x - 6xy^{2} - 8y^{3}
\end{align}$$

Firstly, find the partial derivatives:

$$\begin{align}
f_{x} &= 24 - 6y^{2} \\  
f_{y} &= -12xy - 24y^{2} \\
\end{align}$$

Next, find the critcal points by solving the equations:

$$\begin{align}
f_{x} = 24 - 6y^{2} &= 0 \\
y &=\pm 2 \\
\\
Substituting\ the\ value\ of\ y=\pm2\ into: \\
f_{y} = -12xy - 24y^{2} &= 0 \\
\\
for\ y = -2:\ \\
-12(-2)x - 24(-2)^{2} &= 0 \\
x &= 4 \\
\\
for\ y = 2:\ \\
-12(2)x - 24(2)^{2} &= 0 \\
x &= -4 \\
\end{align}$$

Therefore, the points are (-4, 2) and (4, -2). Moreover, at (-4,2), $z = 24(-4) - 6(-4)2^2 -8(2)^3 = -64$ and at (4,-2), $z = 24(4) - 6(4)(-2)^2 -8(-2)^3 = 64$. So the two critical points are (-4,2,-64) and (4,-2,64).

Now, using the second derivative, $D(a,b) = f_{xx}(a,b) f_{yy}(a,b)-[f_{xy}(a,b)]^{2}$:
$$\begin{align}
f_{xx} &= 0 \\  
f_{yy} &= -12x - 48y \\
f_{xy} &= -12y \\
\\
D &= 0 * (-12x - 48y) - [-12y]^{2} \\
D &= -144y^{2} \\
\end{align}$$

At (-4, 2, -64), D = -576, and at (4,-2, 64), D = -576, therefore both are saddle points.

***
### Problem 3  

A grocery store sells two brands of a product, the "house" brand and a "name" brand. The manager estimates that if she sells the "house" brand for x dollars and the "name" brand for y dollars, she will be able to sell $81 - 21x + 17y$ units of the "house" brand and $40 + 11x - 23y$ units of the "name" brand.

Find the revenue function R(x,y).

$$\begin{align}
R(x,y) &= (81 - 21x + 17y)x + (40 + 11x - 23y)y \\
&= (81x - 21x^{2} + 17xy) + (40y + 11xy - 23y^{2}) \\
&= 81x - 21x^{2} + 28yx + 40y - 23y^{2}
\end{align}$$

What is the revenue if she sells the "house" brand for \$2.30 and the "name" brand for \$4.10?

```{r}
x = 2.3
y = 4.1

# Revenue =
81*x - 21*(x^2) + 28*y*x + 40*y - 23*(y^2)
```

***
### Problem 4  

A company has a plant in Los Angeles and a plant in Denver. The firm is committed to produce a total of 96 units of a product each week. The total weekly cost is given by $C(x, y) = \frac{1}{6}x^{2} + \frac{1}{6}y^{2} + 7x + 25y + 700$, where x is the number of units produced in Los Angeles and y is the number of units produced in Denver. How many units should be produced in each plant to minimize the total weekly cost?

$$\begin{align} 
x + y &= 96 \\
y &= 96 - x \\
Substituting\ the\ value\ of\ y=96 - x\ into:\\
C(x, y) &= \frac{1}{6}x^{2} + \frac{1}{6}y^{2} + 7x + 25y + 700 \\
C(x, y = 96 - x) &= \frac{1}{6}x^{2} + \frac{1}{6}(96 - x)^{2} + 7x + 25(96 - x) + 700 \\
&= \frac{1}{6}x^2 + 1536 -32x +\frac{1}{6}x^2 + 7x + 2400 -25x + 700 \\
&= \frac{1}{3}x^2 -50x + 4636 \\
\end{align}$$

Now find the derivative and evaluate to find the critical point:
$$\begin{align} 
C^{(1)}(x, y = 96 - x) &=  \frac{2}{3}x -50 = 0 \\
x &= 75 \\
And\ now,\ y &= 21 \\
\end{align}$$

Thus the company will need to produce 75 units in Los Angeles and 21 units in Denver.

***
### Problem 5  

Evaluate the double integral on the given region.

$$\begin{align} 
\iint_R (e^{8x+3y}) dA, \quad R:2 \leq x \leq 4 \ \&\ 2 \leq y \leq 4
\end{align}$$

Write your answer in exact form without decimals.

$$\begin{align}
\iint_R (e^{8x+3y}) dA &= \int_2^4 \left[ \int_2^4 (e^{8x+3y}) dx \right] \ dy \\
&= \int_2^4 \left[ \frac{1}{8} \int_2^4 (e^{8x+3y}) \  8 \ dx \right] \ dy \\
&= \int_2^4 \left[ \left. \frac{1}{8} (e^{8x+3y}) \right|_2^4 \right] \ dy \\
&= \int_2^4 \left[ \frac{1}{8} \left[ (e^{8(4)+3y}) - (e^{8(2)+3y})\right] \right] \ dy \\
&= \int_2^4 \left[ \frac{1}{8} (e^{32}e^{3y}) - (e^{16} e^{3y}) \right] \ dy \\
&= \int_2^4 \left[ \frac{1}{8} (e^{32} - e^{16}) e^{3y} \right] \ dy \\
&= \frac{1}{8} (e^{32} - e^{16}) \int_2^4 \frac{1}{3} \left[ e^{3y} \right] 3 \ dy \\
&= \frac{1}{8} (e^{32} - e^{16}) \frac{1}{3} \left[ \left. e^{3y} \right|_2^4 \right]\\
&= \frac{1}{8} (e^{32} - e^{16}) \frac{1}{3} \left( e^{3(4)} - e^{3(2)} \right) \\
&= \frac{1}{24} (e^{32} - e^{16}) \left( e^{12} - e^6 \right) \\
&= \frac{1}{24} (e^{44} - e^{38} - e^{28} + e^{22}) 
\end{align}$$

